Discrete Quantum Theories and Computing

### 2018-11-19 dissertation.tex
dissertation.lyx "2.2 Quantum Probability" Updating...

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events \cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}.
Given a finite sample space~$\Omega$ representing all possible outcomes,
and its power set $2^{\Omega}$ as the classical event space, a classical
probability measure~$\mu$ maps an event~$E$ to a number $\mu\left(E\right)\in\left[0,1\right]$
specifying how likely one of the outcome in $E$ actually happens.
To maintain the coherence, we want $\mu$ subject to the following
constraints. When $E$ contains no outcome, $E$ is the least likely
to happen, i.e., $\mu\left(\emptyset\right)=0$. When $E$ contains
all outcomes, $E$ is the most likely to happen, i.e., $\mu\left(\Omega\right)=1$.
If we know the likelihood of an event, $\mu\left(E\right)$, it is
natural to assume the likelihood of the event’s complement is just
the opposite of $\mu\left(E\right)$, i.e., $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$.
For each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$,
the likelihood of the union event is additive, $\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$.

The previous abstraction

>>>>

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events \cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}.
Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps an event~$E$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcome in $E$ actually happens. To maintain the
coherence, we want $\mu$ subject to the following constraints. When
$E$ contains no outcome, $E$ is the least likely to happen, i.e.,
$\mu\left(\emptyset\right)=0$. When $E$ contains all outcomes, $E$
is the most likely to happen, i.e., $\mu\left(\Omega\right)=1$. If
we know the likelihood of an event, $\mu\left(E\right)$, it is natural
to assume the likelihood of the event’s complement is just the opposite
of $\mu\left(E\right)$, i.e., $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$.
For each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$,
the likelihood of the union event is additive, $\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$.

Since the previous abstraction disregards the underlying process generating
the outcomes, this process might be a quantum measurement. Let us
prepare a beam of a kind of spin~$1$ particles. In principle, we
can preform a Stern-Gerlach type experiment to splits the beam according
to the eigenvalues of the observable $\mathbf{O}_{0}=0\,\proj{0}+1\,\proj{1}+2\,\proj{2}$
\cite{:/content/aip/journal/jmp/21/1/10.1063/1.524312,peres1995quantum}.
If we only send one particle instead of a beam, it could be in three
possible states after the experiment, i.e., the outcome of this process
is $\Omega_{0}=\left\{ \ket{0},\ket{1},\ket{2}\right\} $. All possible
events are $\emptyset$, $\left\{ \ket{0}\right\} $, $\left\{ \ket{1}\right\} $,
$\left\{ \ket{2}\right\} $, $\left\{ \ket{0},\ket{1}\right\} $,
$\left\{ \ket{0},\ket{2}\right\} $, $\left\{ \ket{1},\ket{2}\right\} $,
and $\Omega$. Given any particle, it specifies the likelihood of
each event $\mu_{0}\left(E\right)$, and the function $\mu_{0}:2^{\Omega}\rightarrow\left[0,1\right]$
is a probability measure satisfying the conditions listed in previous
paragraph.

The different of the quantum world is that the probability of previous
experiment is correlated to the other experiment if we use the same
particle, but change the beam splitter. Consider a beam splitter corresponding
to the observable $\mathbf{O}_{1}=0\,\proj{\ps}+1\,\proj{\ms}+2\,\proj{2}$,
where $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$ and $\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}$.
The outcome of this process is $\Omega_{1}=\left\{ \ket{\ps},\ket{\ms},\ket{2}\right\} $.
All possible events are $\emptyset$, $\left\{ \ket{\ps}\right\} $,
$\left\{ \ket{\ms}\right\} $, $\left\{ \ket{2}\right\} $, $\left\{ \ket{\ps},\ket{\ms}\right\} $,
$\left\{ \ket{\ps},\ket{2}\right\} $, $\left\{ \ket{\ms},\ket{2}\right\} $,
and $\Omega_{1}$. When we can formulate the likelihood of each event
as another probability measure $\mu_{1}:2^{\Omega}\rightarrow\left[0,1\right]$,
we found the likelihood of the same event of these two experiments
are the same if we use the same particle, i.e., $\mu_{0}\left(\emptyset\right)=\mu_{1}\left(\emptyset\right)$
and $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$.

Since the probability is always the same if the particle and the event
are the same no matter which beam splitter is actually used. It is
more natural to define the set of quantum event~$\events$ as all
possible events using different beam splitters. Since we are focus
on the probability instead of outcome, and the probabilities of events
$\left\{ \ket{0},\ket{1}\right\} $ and $\left\{ \ket{\ps},\ket{\ms}\right\} $
are always the same, it is reasonable to identify them as the same
event.

, and consider the quantum probability measure $\mu:\events\rightarrow\left[0,1\right]$
as a function assigning the likelihood to every quantum event
